---
title: "MB1A Practical 1 Part 3"
author: "Katy Brown"
format: html
---

P3 Exercise 1

Round the number 3.14159 to the nearest whole number.

```{r}
round(3.14159)
```

P3 Exercise 2

Repeat the number 5, 7 times, first by specifying arguments in order and then using their names.

```{r}
rep(5, 7)
```

```{r}
rep(x=5, times=7)
```

P3 Exercise 4
Create character, numerical and logical variables, print them, then check their classes.

```{r}
charvar = "hello"
numvar = 5
logvar = TRUE

print (charvar)
print (class(charvar))

print (numvar)
print (class(numvar))

print (logvar)
print (class(logvar))
```

P3 Exercise 5

Create variables of different types showing attributes of cells.

```{r}
cell_id_1 <- "CX1"
cell_id_2 <- "CX2"
treated_1 <- TRUE
treated_2 <- FALSE
num_cells_before_1 <- 5415
num_cells_before_2 <- 8123
num_cells_after_1 <- 1432
num_cells_after_2 <- 7543

```

P4 Exercise 1 

Combine variables above into vectors

```{r}
cell_ids <- c(cell_id_1, cell_id_2)
treated <- c(treated_1, treated_2)
cells_before <- c(num_cells_before_1, num_cells_before_2)
cells_after <- c(num_cells_after_1, num_cells_after_2) 
```

P4 Exercise 2
Isolate value 4 in this vector
```{r}
all_my_pets <- c("dog", "horse", "guinea pig", "chicken", "cat")

val_4 <- all_my_pets[4]

print(val_4)
```
