---
title: "MB1A Practical 1 Part 3"
author: "Katy Brown"
format: html
---

Exercise 1

Round the number 3.14159 to the nearest whole number.

```{r}
round(3.14159)
```

Exercise 2

Write some code to repeat the number 5, 7 times, first by specifying arguments in order and then using their names.

```{r}
rep(5, 7)
```

```{r}
rep(x=5, times=7)
```

Exercise 4
Create character, numerical and logical variables, print them, then check their classes.

```{r}
charvar = "hello"
numvar = 5
logvar = TRUE

print (charvar)
print (class(charvar))

print (numvar)
print (class(numvar))

print (logvar)
print (class(logvar))
```

Exercise 5

You have conducted an experiment in which you have treated one batch of cells with radiation and kept the other as a control, then counted the number of live cells before and after.

The attributes of your two batches of cells are in this table:
cell_ID 	treated 	num_cells_before 	num_cells_after
CX1 	TRUE 	5415 	1432
CX2 	FALSE 	8123 	7543

```{r}
cell_id_1 <- "CX1"
cell_id_2 <- "CX2"
treated_1 <- TRUE
treated_2 <- FALSE
num_cells_before_1 <- 5415
num_cells_before_2 <- 8123
num_cells_after_1 <- 1432
num_cells_after_2 <- 7543

```