---
title: "Worked Example - Bird Migration"
author: "Katy Brown"
format: html
---
```{r}
library(tidyverse)
library(gt)
library(maps)
```

Read the bird breeding site data frame and look at it with `gt()`.

```{r}
bird_breeding_sites <- read_csv("data/breeding_site_locations.csv")

gt(bird_breeding_sites)
```

Preview the data with `glimpse()`.

```{r}
glimpse(bird_breeding_sites)
```

Read the bird migration site data.
```{r}
bird_migration_sites <- read_csv("data/migratory_connectivity_dataset.csv")
gt(head(bird_migration_sites))
```
Summarise the data frame

```{r}
glimpse(bird_migration_sites)
```

Filter columns in bird_breeding_sites data frame

```{r}
bird_breeding_selected <- bird_breeding_sites |>
  select(Site, Longitude, Latitude) |>
  distinct() |>
  rename(Breeding_Longitude = Longitude, Breeding_Latitude = Latitude)

gt(bird_breeding_selected)
```

Filter columns in bird_migration_sites data frame

```{r}
bird_migration_selected <- bird_migration_sites |>
  select(
    Site,
    winterLongitude,
    winterLatitude,
    pslon,
    pslat,
    iblon,
    iblat,
    awestlon,
    awestlat
  ) |>
  unique() |>
  rename(
    Winter_Longitude = winterLongitude,
    Winter_Latitude = winterLatitude,
    Stopover_Longitude = pslon,
    Stopover_Latitude = pslat,
    Iberian_Longitude = iblon,
    Iberian_Latitude = iblat,
    Westmost_Longitude = awestlon,
    Westmost_Latitude = awestlat
  )

gt(head(bird_migration_selected))
```



```{r}
bird_breeding_exc <- bird_breeding_selected |>
  filter(Site != "Spain")

bird_migration_exc <- bird_migration_selected |>
  filter(Site != "Spain")

gt(bird_breeding_exc)
```
Calculate median longitude and latitude in the bird_migration data frame.

```{r}
bird_migration_grouped <- bird_migration_exc |> group_by(Site)

bird_migration_summary <- bird_migration_grouped |>
  summarise(
    Stopover_Longitude = median(Stopover_Longitude, na.rm = TRUE),
    Stopover_Latitude = median(Stopover_Latitude, na.rm = TRUE),
    Iberian_Longitude = median(Iberian_Longitude, na.rm = TRUE),
    Iberian_Latitude = median(Iberian_Latitude, na.rm = TRUE),
    Westmost_Longitude = median(Westmost_Longitude, na.rm = TRUE),
    Westmost_Latitude = median(Westmost_Latitude, na.rm = TRUE),
    Winter_Longitude = median(Winter_Longitude, na.rm = TRUE),
    Winter_Latitude = median(Winter_Latitude, na.rm = TRUE),
  )
gt(bird_migration_summary)
```

Join the two data frames

```{r}
combined_bird_data <- full_join(bird_breeding_exc, bird_migration_summary)

gt(combined_bird_data)
```
```{r}
combined_bird_piv <- combined_bird_data |>
  pivot_longer(
    cols = ends_with(c("_Longitude", "_Latitude")),
    names_sep = "_",
    names_to = c("Location_Type", ".value")
  )

gt(combined_bird_piv)
```

Split into spring and autumn

```{r}
spring_vals <- c("Breeding", "Stopover", "Winter")
autumn_vals <- c("Breeding", "Iberian", "Westmost", "Winter")

spring_data <- combined_bird_piv |>
  filter(Location_Type %in% spring_vals)

autumn_data <- combined_bird_piv |>
  filter(Location_Type %in% autumn_vals)
```

Save file

```{r}
write_csv(spring_data, "output/birds_spring.csv")
write_csv(autumn_data, "output/birds_autumn.csv")
```

Calculate axis limits

```{r}
min_lon <- spring_data |> pull(Longitude) |> min()
max_lon <- spring_data |> pull(Longitude) |> max()
min_lat <- spring_data |> pull(Latitude) |> min()
max_lat <- spring_data |> pull(Latitude) |> max()
```

Plot the data
```{r}
ggplot(
  spring_data,
  aes(x = Longitude, y = Latitude, color = Site, shape = Location_Type)
) +
  geom_point(size = 4) +
  annotation_borders(linewidth = 0.2) +
  coord_quickmap(
    xlim = c(min_lon - 2, max_lon + 2),
    ylim = c(min_lat - 2, max_lat + 2)
  ) +
  geom_path(aes(group = Site), size = 1) +
  scale_color_brewer(palette = "Dark2") +
  theme_bw() +
  ggtitle("Spring Migration Routes")
```

```{r}
ggplot(
  autumn_data,
  aes(x = Longitude, y = Latitude, color = Site, shape = Location_Type)
) +
  geom_point(size = 4) +
  annotation_borders(linewidth = 0.2) +
  coord_quickmap(
    xlim = c(min_lon - 2, max_lon + 2),
    ylim = c(min_lat - 2, max_lat + 2)
  ) +
  geom_path(aes(group = Site), size = 1) +
  scale_color_brewer(palette = "Dark2") +
  theme_bw() +
  ggtitle("Autumn Migration Routes")
```